A Computational Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-04-01 paper/paper.tex
stuff

\begin{abstract}
The default mode network (DMN) is the brain manifestation of the human
baseline mental activity.
%
Many research streams agree on its likely role in evolutionarily adaptive
envisioning of mental scenarios to predict the environment.
The DMN would hence be dedicated to continuous autobiographical memory retrieval,
generation of hypothetical outcomes,
and reward contingency evaluation when letting the mind go.
It explains its highest energy consumption in the brain and
its intimate coupling with conscious awareness.
%
This concept paper proposes a process model that describes
\textit{how} the DMN may actually implement continuous
environmental assessment and prediction to guide action choices.
DMN function is recast in mathematical terms
from the perspective of
% control theory and
reinforcement learning.
We argue that our formal account of DMN function
explains existing cognitive accounts as special cases and
parsimoneously explains a variety of experimental findings
in animals and humans.
%
Formal models of the computational processes subserved in the DMN
could offer access routes to the statistical regularities
of human predictive behavior.
\end{abstract}

@@

Computing an optimal policy then corresponds to finding the parameters $\theta^Q_{k+1}$ which minimize the following mean-squared loss function
\begin{equation}
  \mathcal L(\theta^Q_{k+1})
  % = \mathbb E_{(s, a, r, s') \sim \mathcal U(D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{k+1}))^2]
  = \mathbb E_{(s, a, r, s') \sim \mathcal U(D)}\left[\frac{1}{2}(Q(s, a|\theta^Q_{k+1}) - y)^2\right],
  \label{eq:oracle}
\end{equation}
where
$ y: = r + \gamma \max_{a' \in \mathcal A} Q(s', a'|\theta^Q_k) = r + \gamma Q(s', \pi(s')|\theta^Q_k)$ is the prediction target.
% While $y_t$ is also dependent on $\theta^Q$, this is typically ignored.
For instance, a general linear model with a kernel $\phi$ would be of the
form
$${Q}(s, a|\theta^Q) = \phi(s,a)^T\theta^Q.$$
$\phi(s,a)$ would represent a high-level hand-crafted representation in an
appropriate form of the state-action pairs $(s,a)$,
as proposed by \cite{songNIPS2016}.
%  and Parr et al. 2003 (JMLR)].
A recently proposed breakthrough alternative
\citep{mnih2015,silver2016mastering}, is to learn this
representation using a deep neural net, leading to the so-called \textit{Deep Q-learning} family of methods which
are by far the current state-of-the-art in RL. In such a case, $\theta^Q$
would correspond to the strengths of low-level synaptic or high-level functional connections between different elements
in brain networks.
\textbf{XXX perhaps add one sentence to rephrase and detail the idea}. \textbf{XXX Elvis: My biology is very poor.
  The idea of the statement was to say where the parameters $\theta^Q$ would ``live'' in the brain.
  Would these be synaptic strengths ? effective functional connectivity ?
If you know this, please could you rephrase ?}

>>>>

\begin{abstract}
The default mode network (DMN) is the brain manifestation of the human
baseline mental activity.
%
Many research streams agree on its likely role in evolutionarily adaptive
envisioning of mental scenarios to predict the environment.
The DMN would hence be dedicated to continuous autobiographical memory retrieval,
generation of hypothetical outcomes,
and reward contingency evaluation when letting the mind go.
It explains its highest energy consumption in the brain and
its intimate coupling with conscious awareness.
%
This concept paper proposes a process model that describes
\textit{how} the DMN may actually implement continuous
environmental assessment and prediction to guide action choices.
DMN function is recast in the mathematical terms of
reinforcement learning and control by means of Markov decision processes.
We argue that our formal account of DMN function
explains existing cognitive accounts as special cases and
parsimoneously explains a variety of experimental findings
in animals and humans.
%
Formal models of the computational processes subserved in the DMN
could offer access routes to the statistical regularities
of human predictive behavior.
\end{abstract}

@@

Computing an optimal policy then corresponds to finding the parameters $\theta^Q_{k+1}$ which minimize the following mean-squared loss function
\begin{equation}
  \mathcal L(\theta^Q_{k+1})
  % = \mathbb E_{(s, a, r, s') \sim \mathcal U(D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{k+1}))^2]
  = \mathbb E_{(s, a, r, s') \sim \mathcal U(D)}\left[\frac{1}{2}(Q(s, a|\theta^Q_{k+1}) - y)^2\right],
  \label{eq:oracle}
\end{equation}
where
$ y: = r + \gamma \max_{a' \in \mathcal A} Q(s', a'|\theta^Q_k) = r + \gamma Q(s', \pi(s')|\theta^Q_k)$ is the prediction target.
% While $y_t$ is also dependent on $\theta^Q$, this is typically ignored.
For instance, a general linear model with a kernel $\phi$ would be of the
form
$${Q}(s, a|\theta^Q) = \phi(s,a)^T\theta^Q.$$
$\phi(s,a)$ would represent a high-level hand-crafted representation in an
appropriate form of the state-action pairs $(s,a)$,
as proposed by \cite{songNIPS2016}.
%  and Parr et al. 2003 (JMLR)].
A recently proposed, practically successful alternative
\citep{mnih2015,silver2016mastering} is to learn this
representation using a deep neural network, leading to the so-called \textit{Deep Q-learning} family of methods which
are by far the current state-of-the-art in RL.
The set of $\theta^Q$ parameters that instatiate the nonlinear interactions
between layers of the artifical neural network
may find a neurobiological correspondence in the adaptive strengths of axonal
connections between neurons from the different levels
of the neural processing hierarchy
\citep{mesulam1998sensation, taylor2015global}.
